Hysteresis in human computation:\\ how one task affects another

### 2014-12-17 docs/publications/2014.09_SCIENCE/man.tex
another quick passt

Given the increasing prevalence of microtask-based research methods, there
is a desire to understand the factors affecting the reliability of HPU
outputs.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  
Yet to our knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation. The experiments we 
present here show that, unlike CPUs and GPUs, HPUs are subject to 
\textit{hysteresis}, meaning that their outputs depend on the 
\textit{history} of their inputs.

@@

It is difficult to measure the change in a probability distribution based on
samples alone (\#), so we reformulate the problem slightly. Consider that,  
no matter the ultimate purpose of a microtask study, a microtask response
is used in \textit{some kind} of subsequent decision or 
further processing.
We denote this downstream decision by $\mathcal{D}$.  Since $\mathcal{D}$
takes a microtask response as input, biases in the microtask response will 
in general lead to biases in the output of $\mathcal{D}$.  
We can make use of this fact, 
by placing a machine learning algorithm (a classifier), in the role of 
$\mathcal{D}$.  This classifier takes as input a worker's responses to test 
tasks, from which it learns to infer the worker's treatment 
(i.e. the initial tasks she was given). 
Intuitively, the classifier's ability to infer the worker's treatment
depends on how strongly the exposure to prior tasks influences responses to 
subsequent ones.  Formally, the bias, $\theta$, is bounded 
by the accuracy of the classifier, $\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
Thus, measuring the performance of the classifier allows us to establish a
lower bound on the extent of bias. This approach can be used to bound the 
bias induced by any perturbation, including both intertask effects and 
framing.  Bias ranges from 0 (meaning there are no intertask effects), to 1 
(or 100\%) (meaning that changing initial tasks leads to a completely 
different set of responses to test tasks).

@@

Using a na\"ive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to the results obtained 
if a support vector machine is used.), we found a bias of 30\% between 
workers from the \textit{food} and \textit{objects} treatments, due to 
intertask effects.
As a point of comparison, we performed a similar experiment in which we 
framed the purpose of the tasks.  Workers were either told that the tasks 
were ``Funded by the laboratory for the visual perception of Food and 
Ingredients'', or ``\ldots of Objects and Tools''.  We estimated the bias 
induced by framing using a na\"ive Bayes classifier, but did not find a 
statistically significant effect (see Fig.~\ref{fig:theta}).  This suggests
that prior task exposure is a stronger priming modality than framing.

@@

It was only when we combined the use of framing, with an active 
reiteration step, that bias reached a level comparable to that achieved 
by intertask effects.  In the experiment \textit{echo-food-objects},
after framing the purpose of the task around either the recognition of food
or objects, we required the worker to echo back the purpose of the task
using combo-box input.  In this case, workers exposed to different 
``echoed framing'' showed a bias of about 35\% in the labels provided during 
test tasks
(see Fig.~\ref{fig:theta}A). However, it is possible that by requiring 
workers to reiterate the language used in framing, we provided an implicit 
but nonetheless strong signal that the worker must act on the framing 
statement.  In this sense, \textit{echo-food-culture} may reflect the 
influence of an instruction rather than framing.

@@

To better understand the nature of the 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, within
a given experiment, the food-primed workers would use more food-related words 
than their non-food-primed counterparts.  However, this was not generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks
(\# food related words were identified using the wordnet corpus, augmented
with additional words obtained by crawling a recipe website (see supplementary
information for details)) (see Fig.~\ref{fig:specificity}).  This finding
rules out a simplistic conception of how intertask effects guide the worker's
attention.  It seems that when workers encounter content in 
prior tasks, the probability that they will refer to similar content in 
test tasks changes significantly, although it does not necessarily increase.

@@

This suggests that priming, by prior task exposure as well as framing,
might have led workers to use more refined or specialized words to refer to
the concept present in the prime.
To test this directly, we used the wordnet corpus to operationalize the 
notion of word specialization.
Wordnet provides a set of hypernym-hyponym relations between \# English 
nouns.  
Hypernyms are generalizations (for example, ``bread'' is a hypernym for 
``pumpernickel''), while hyponyms are specializations.
To assess the relative specialization of words from two different treatments,
we compared all pairs of words from one treatment to the other. 
We tallied the cases where a word from one treatment was more specialized
than one from the other, as well as the cases with roles reversed.  
The difference of these tallies, as a percentage of their total, gave the
relative specialization of one treatments vocabulary over the other
(\#) (see Fig.~\ref{fig:specificity}).  
Using this approach, we compared the specialization of food-related words 
for the treatments of all experiments.

In all experiments except \textit{frame-food-culture}, food-primed workers 
used significantly more specialized words than their non-food-primed 
counterparts (about 15\% more; see Fig.~\ref{fig:specificity}C).
It is interesting that such substantial increases in both the lexical 
richness and specialization of food-related words among food-primed workers
held for \textit{intertask-food-culture}, where, again, we observed fewer 
references to food by food-primed workers overall. These observations point 
to countervailing factors: one tending to activate the more specialized and 
less common food-related words (yielding greater lexical richness and 
specialization), and the other tending to suppress certain, presumably more 
common or generic words (yielding fewer food-related words in total).

@@

Prior to this study, little thought appears to have been given to the 
ordering and bundling of tasks.  But our findings show that these are 
powerful design factors.  A natural response to our findings 
might be simply to randomize task ordering, a practice that is commonly 
observed.  But the sheer extent of bias that we observed leads us to question 
whether that is the best approach, since this would tend to raise 
the ``noise floor'' of the study.

Furthermore, our findings suggest that, through careful task engineering,
it might be able to achieve much more.  The boost that we observed to the 
specialization and richness of workers' responses motivates further 
investigation to acquire a deeper understanding of intertask effects.  
We hope that future work will yield techniques to leverage intertask effects, 
enabling one to tune the focus, diversity, and specialization of worker 
responses.

A consistent goal in human computation is the achievement of expert-level
judgments from non-expert workers.  This has been achieved in some
applications (\#). The distinction between experts and novices can be 
attributed in part to the knowledge of facts and 
heuristics. But to a large degree, this distinction lies in the ability of 
experts to direct their focus towards salient features, while filtering out 
non-salient ones \cite{kellman2009perceptual}.  Using prior task exposure to 
guide focus and salience attribution might well enable expert-level 
judgment in a wider variety of crowdsourced applications.

Here we have shown that a large bias can be introduced into microtask 
responses, simply from the performance of earlier tasks. 
These effects are much stronger than framing, and are only comparable when
framing is combined with a requirement that the worker actively echo
the framing language.  While our findings raise serious concerns about the 
current state of microtask design, it also reveals opportunities for more 
refined control over worker focus and acuity.  Any application that relies
on the repetitive use of human judgment is likely subject to the phenomena
we described here.  It remains to be determined whether intertask effects 
can be leveraged to increase the performance and reliability of human 
computation.

>>>>

Given the increasing prevalence of microtask-based research methods, there
is a desire to understand factors affecting the reliability of HPU
outputs.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  
Yet to our knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation. The experiments we 
present here show that, unlike CPUs and GPUs, HPUs are subject to 
\textit{hysteresis}, meaning that their outputs depend on the 
\textit{history} of their inputs.

@@

It is difficult to measure the change in a probability distribution based on
samples alone (\#), so we reformulate the problem slightly. Consider that,  
no matter the ultimate purpose of a microtask study, a microtask response
is used in \textit{some kind} of subsequent decision or 
further processing.
We denote this downstream decision by $\mathcal{D}$.  Since $\mathcal{D}$
takes a microtask response as input, biases in the microtask response will 
in general lead to biases in the output of $\mathcal{D}$.  
We can make use of this fact, to estimate the bias, 
by placing a machine learning algorithm (a classifier), in the role of 
$\mathcal{D}$.  This classifier takes as input a worker's responses to test 
tasks, from which it learns to infer the worker's treatment 
(i.e. the initial tasks she was given). 
Intuitively, the classifier's ability to infer worker treatments
depends on how strongly prior tasks affect responses to 
subsequent ones.  Formally, the bias, $\theta$, is bounded 
by the accuracy of the classifier, $\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
Thus, measuring the performance of the classifier allows us to establish a
lower bound on the extent of bias. This approach can be used to bound the 
bias induced by any perturbation, including both intertask effects and 
framing.  Bias ranges from 0 (meaning there are no intertask effects), to 1 
(or 100\%) (meaning that the change of initial tasks induced a completely 
different set of responses to test tasks).

@@

Using a na\"ive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to the results obtained 
if a support vector machine is used.), we found a bias of 30\% between 
workers from the \textit{food} and \textit{objects} treatments, due to 
intertask effects.

As a point of comparison, we performed a similar experiment in which we 
framed the purpose of the tasks.  Workers were either told that the tasks 
were ``Funded by the laboratory for the visual perception of Food and 
Ingredients'', or ``\ldots of Objects and Tools''.  We did not observe a 
statistically significant framing-induced bias, using a na\"ive Bayes 
classifier (see Fig.~\ref{fig:theta}).  This suggests
that prior task exposure is a stronger priming modality than framing.

@@

It was only when we combined the use of framing, with an active 
reiteration step, that bias reached a level comparable to that achieved 
by intertask effects.  In the experiment \textit{echo-food-objects},
after framing the purpose of the task around either the recognition of food
or objects, we required the worker to echo back the purpose of the task
using a combo-box input.  In this case, workers exposed to different 
``echoed framing'' showed a bias of about 35\% in the labels provided during 
test tasks
(see Fig.~\ref{fig:theta}A). However, it is possible that by requiring 
workers to echo the language used in framing, we provided an implicit 
but nonetheless strong signal of our intent that workers act on the 
framing statement.  In this sense, \textit{echo-food-culture} may reflect the 
influence of an \textit{instruction} rather than framing.

@@

To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, within
a given experiment, the food-primed workers would use more food-related words 
than their non-food-primed counterparts.  However, this was not generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks
(\# food related words were identified using the wordnet corpus, augmented
with additional words obtained by crawling a recipe website (see supplementary
information for details)) (see Fig.~\ref{fig:specificity}).  This finding
rules out a simplistic conception of how intertask effects guide the worker's
attention.  It seems that when workers encounter content in 
prior tasks, the probability that they will refer to similar content in 
subsequent tasks changes significantly, although it does not necessarily 
increase.

@@

This lexical enrichment suggests that priming, whether by prior task 
exposure or framing, leads workers to use more refined 
or specialized words, in reference to concepts present in the prime.
To test this directly, we used the wordnet corpus to operationalize the 
notion of word specialization.
Wordnet provides a set of hypernym-hyponym relations between \# English 
nouns.  
Hypernyms are generalizations (for example, ``bread'' is a hypernym for 
``pumpernickel''), while hyponyms are specializations.
To assess the relative specialization of words from two different treatments,
we compared all pairs of words from one treatment to the other. 
We tallied the cases where one treatment's word was more specialized
than the other, as well as the cases wher roles were reversed.  
The difference of these tallies, as a percentage of their total, served as 
the relative specialization of one treatment's vocabulary relative to the 
other (\#) (see Fig.~\ref{fig:specificity}).  
Using this approach, we compared the specialization of food-related words 
for the treatments of all experiments.

In all experiments except \textit{frame-food-culture}, food-primed workers 
used significantly more specialized words than their non-food-primed 
counterparts (about 15\% more; see Fig.~\ref{fig:specificity}C).
It is interesting that such substantial increases in both the lexical 
richness and specialization of food-related words 
held for \textit{intertask-food-culture}, where, again, we observed fewer 
references to food by food-primed workers overall. These observations point 
to countervailing factors: one factor tending to activate the more 
specialized and less common food-related words 
(yielding greater lexical richness and specialization), and the other tending 
to suppress certain, presumably more common or generic words 
(yielding fewer food-related words in total).

@@

Prior to this study, little thought appears to have been given to the 
ordering and bundling of microtasks.  But our findings show that these are 
powerful design factors.  A natural response to our findings 
might be simply to randomize task ordering, a practice that is commonly 
observed.  But the sheer extent of bias that we observed leads us to question 
whether that is the best approach, since this would tend to raise 
the ``noise floor'' of the study.

Furthermore, our findings suggest that, through careful task engineering,
it might be possible to achieve much more.  The boost that we observed to the 
richness and specialization of workers' responses motivates further 
investigation to acquire a deeper understanding of intertask effects.  
We hope that future work will yield techniques to leverage intertask effects, 
enabling one to tune the focus, diversity, and specificity of worker 
responses.

A consistent goal in human computation is the achievement of expert-level
judgments from non-expert workers (\#).  This has been achieved in some
applications (\#). The distinction between experts and novices can be 
attributed in part to the posession of specialized knowledge and heuristics. 
But to a large degree, this distinction lies in the ability of 
experts to direct their focus toward salient features, while filtering out 
non-salient ones \cite{kellman2009perceptual}.  Using prior task exposure to 
guide focus and salience attribution might enable expert-level 
judgment in a wider variety of crowdsourced applications.

Here we have shown that a large bias can be introduced into microtask 
responses, simply from the performance of earlier tasks. 
These effects are much stronger than framing.  
While our findings raise serious concerns about the 
current state of microtask design, they also reveal potential opportunities 
for more 
refined control over worker focus and acuity.  Any application that relies
on the repetitive use of human judgment is likely subject to the phenomena
we described here.  It remains to be determined by future work whether 
intertask effects can be leveraged for more performant and reliable human 
computation.
